---
title: "Forecasting Averaging"
author: "Nikesh Patel"
date: "`r format(Sys.Date(),'%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    fig_caption: yes
    theme: united
    toc_float: no
  pdf_document:
    toc: yes
---
$~$

```{r setup, include = FALSE}
rm(list = ls())
graphics.off()
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r Packages, warning=FALSE, include=FALSE}
require(tidyverse)
require(kableExtra)
require(tidyquant)
require(lubridate)
require(timetk)
require(tsibble)
require(ggplot2)
require(reshape2)
require(fpp3)
```
# ** Abstract**

# **Introduction**

We tend to look at inflation to see how much prices will be increasing. Looking at inflation, the United States government can be able to utilize both fiscal and monetary policy. Inflation also effects small businesses and investments too. Our goal is to use to use the following code to forecast averages of inflation. We will use a Phillips curve as a base to evaluate the performance of our models. According to the Phillips Curve, the unemployment rate and the inflation rate should be inverted with each other. That being said as inflation is high, the unemployment rate should be low and vice-versa. We would also want to 

# **Hypothesis**

Given the fact that the Phillips Curve shows that the inflation rate and the unemployment rate is the inverse of each other, the unemployment rate could possibly be the best overall predictor of inflation rate itself. However, one drawback for this variable could be that unemployment rate might not be the best indicator for the shocks from COVID-19.

# **Data Transformations**

First, we want to define the following variables being used:

**"PCEPI"** is the measure of prices people pay for goods and services. This is a good indication of the changes in consumer behavior.

**"UNRATE"** is the annualized unemployment rate. That is the number of unemployed people in comparision to the labor force. 

**"MICH"** is a survey from consumers explaining the median expected price change.

**"TCU"** is the index of total capacity utilization, which is the percentage of resources used by firms/corporations to produce goods. In essence, how much capacity is being used to produced demanded goods.

**"INDPRO"** is the index of total industrial production which measures all output in the United States regarding the sector of manufacturing, mining, electric, etc.  

We will be using the following variables to help see if there is any improvement within forecasting the model. All the data provided is from January, 1982 to current times. Some of the variables needed to be transposed in order to account for seasonality. For example, the variable "MICH" needed to be differenced because the original data was not seasonally adjusted. 


```{r GetData, message=FALSE}
varList <- c("PCEPI", "UNRATE", "MICH", "TCU", "INDPRO")	
X <-
  tq_get(varList, get = "economic.data", from = "1982-01-01") %>%
  mutate(Month = yearmonth(date), value = price) %>%
  dplyr::select(-c(date, price)) %>%
  as_tsibble(index = Month, key = symbol)
Xw <- X %>%
  pivot_wider(names_from = symbol, values_from = value) %>%
  as_tsibble()  %>% 
  drop_na()

Xw <- Xw %>%
mutate(mich = difference(MICH))

Xy <- Xw %>% 
  features(mich, unitroot_kpss) 
```
```{r Mutate, message= FALSE}
Z <- Xw %>% select(c(PCEPI, UNRATE, MICH, TCU, INDPRO)) %>%
  mutate(infl = 1200*log(PCEPI/lag(PCEPI))) %>% 
  mutate(dinfl = infl - lag(infl,1)) %>% 
  mutate(dinfl12 = 100*log(PCEPI/lag(PCEPI,12)) - lag(infl,12)) %>% 
  mutate(unrate = UNRATE - lag(UNRATE)) %>% 
  mutate(tcu = TCU - lag(TCU)) %>% 
  mutate(indpro = 1200*log(INDPRO/lag(INDPRO))) %>% 
  mutate(mich = difference(MICH)) %>%
  select(-c(PCEPI, UNRATE, MICH, TCU, INDPRO)) %>% 
  drop_na()
train_data <- Z %>% filter_index(~ "2019-12")
test_data <- Z %>% filter_index("2020-01" ~ .)
```
```{r Phillips Curve, message= FALSE}
Zm <- melt(Z, "Month")
ggplot(Zm, aes(Month, value)) + 
  geom_line() + 
  facet_wrap(~variable, scales = "free", ncol = 2)
```
After the proper transformations
```{r lags, message= FALSE}
fitPC <- train_data %>% 
  model(
    mUR = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(unrate,12) + lag(unrate,13) + lag(unrate,14) +
                 lag(unrate,15) + lag(unrate,16) + lag(unrate,17) +
                 lag(unrate,18) + lag(unrate,19) + lag(unrate,20) +
                 lag(unrate,21) + lag(unrate,22) + lag(unrate,23)), 
                 
    mtcu = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(tcu,12) + lag(tcu,13) + lag(tcu,14) +
                 lag(tcu,15) + lag(tcu,16) + lag(tcu,17) +
                 lag(tcu,18) + lag(tcu,19) + lag(tcu,20) +
                 lag(tcu,21) + lag(tcu,22) + lag(tcu,23)),
                 
    mindpro = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(indpro,12) + lag(indpro,13) + lag(indpro,14) +
                 lag(indpro,15) + lag(indpro,16) + lag(indpro,17) +
                 lag(indpro,18) + lag(indpro,19) + lag(indpro,20) +
                 lag(indpro,21) + lag(indpro,22) + lag(indpro,23)),
                 
    mmich = TSLM(dinfl12 ~ 1 +
                 lag(dinfl,12) + lag(dinfl,13) + lag(dinfl,14) +
                 lag(dinfl,15) + lag(dinfl,16) + lag(dinfl,17) +
                 lag(dinfl,18) + lag(dinfl,19) + lag(dinfl,20) +
                 lag(dinfl,21) + lag(dinfl,22) + lag(dinfl,23) +
                 lag(mich,12) + lag(mich,13) + lag(mich,14) +
                 lag(mich,15) + lag(mich,16) + lag(mich,17) +
                 lag(mich,18) + lag(mich,19) + lag(mich,20) +
                 lag(mich,21) + lag(mich,22) + lag(mich,23))
                 )
```   

```{r tidy, include=FALSE} 
tidy(fitPC)
```

# **Analysis**


```{r model accuracy}
Fit <- fitPC %>% 
  mutate(Ensemble = (mUR + mtcu + mindpro + mmich)/4)
Fit_check <- accuracy(Fit) %>%
  select(c(".model", ".type", "MAPE")) %>%
  kable(format = "html", table.attr = "style='width:30%;' ") %>%
    kableExtra::kable_styling()

Fit_check

Fit_forecast <- Fit %>% 
  forecast(new_data = test_data)

Fit_accuracy <- accuracy(Fit_forecast, Z) %>%
  select(c(".model", ".type", "MAPE")) %>%
  kable(format = "html", table.attr = "style='width:30%;' ") %>%
    kableExtra::kable_styling()
  
Fit_accuracy
```

```{r plot forecast}
Fit_forecast %>% autoplot(filter(Z, year(Month) > 2016), level = c(95))
```

# **Conclusion**

After further analysis, the initial hypothesis that unemployment rate would be a good indicator of inflation due to the inverse relationship proves to be false. It turns out that TCU (index of total capacity utilization) would be a better model for predicting inflation in the short run (1-6 months). According to the forecast plot, it would seem that the total capacity utilization predicts inflation quite well. However, after 6 months, total capacity utilization would not be the best option. 

In theory, this makes sense due to the fact that total capacity utilization is the percent of resources firms/corporations use to produce goods. During the COVID period (2020-), in the month of January 2021, the total capacity utilization was predicted to be at the same rate of the predicted inflation rate. In this period of time, the economy was above its natural capacity utilization rate, indicating that the demand for goods was really high and supply is trying to catch up. This would indicate that inflation would also increase too. In the long run, different causal variables would create the notion that capacity utilization rate would not be the best predictor of the inflation rate.  

Overall, the TCU variable would be the better option in encapsulating the shocks that occur compared to the other variables when predicting inflation rates. 